### 2016-10-27 Alice_Authorlist_2015-Feb-18.tex
first commit v1

% Collaboration: CERN-LHC-ALICE
% Generation Date is 2015/Feb/18

% How to use:
%%%%%%%%% appendix with author list
%\appendix
%\section{The ALICE Collaboration}
%\label{app:collab}
%\input{authors-list.tex}  %%%%%%% get the latest version before submitting

\begingroup
\small
\begin{flushleft}
J.~Adam\Irefn{org39}\And
D.~Adamov\'{a}\Irefn{org82}\And
M.M.~Aggarwal\Irefn{org86}\And
G.~Aglieri Rinella\Irefn{org36}\And
M.~Agnello\Irefn{org110}\And
N.~Agrawal\Irefn{org47}\And
Z.~Ahammed\Irefn{org130}\And
S.U.~Ahn\Irefn{org67}\And
I.~Aimo\Irefn{org93}\textsuperscript{,}\Irefn{org110}\And
S.~Aiola\Irefn{org135}\And
M.~Ajaz\Irefn{org16}\And
A.~Akindinov\Irefn{org57}\And
S.N.~Alam\Irefn{org130}\And
D.~Aleksandrov\Irefn{org99}\And
B.~Alessandro\Irefn{org110}\And
D.~Alexandre\Irefn{org101}\And
R.~Alfaro Molina\Irefn{org63}\And
A.~Alici\Irefn{org104}\textsuperscript{,}\Irefn{org12}\And
A.~Alkin\Irefn{org3}\And
J.~Alme\Irefn{org37}\And
T.~Alt\Irefn{org42}\And
S.~Altinpinar\Irefn{org18}\And
I.~Altsybeev\Irefn{org129}\And
C.~Alves Garcia Prado\Irefn{org118}\And
C.~Andrei\Irefn{org77}\And
A.~Andronic\Irefn{org96}\And
V.~Anguelov\Irefn{org92}\And
J.~Anielski\Irefn{org53}\And
T.~Anti\v{c}i\'{c}\Irefn{org97}\And
F.~Antinori\Irefn{org107}\And
P.~Antonioli\Irefn{org104}\And
L.~Aphecetche\Irefn{org112}\And
H.~Appelsh\"{a}user\Irefn{org52}\And
S.~Arcelli\Irefn{org28}\And
N.~Armesto\Irefn{org17}\And
R.~Arnaldi\Irefn{org110}\And
T.~Aronsson\Irefn{org135}\And
I.C.~Arsene\Irefn{org22}\And
M.~Arslandok\Irefn{org52}\And
A.~Augustinus\Irefn{org36}\And
R.~Averbeck\Irefn{org96}\And
M.D.~Azmi\Irefn{org19}\And
M.~Bach\Irefn{org42}\And
A.~Badal\`{a}\Irefn{org106}\And
Y.W.~Baek\Irefn{org43}\And
S.~Bagnasco\Irefn{org110}\And
R.~Bailhache\Irefn{org52}\And
R.~Bala\Irefn{org89}\And
A.~Baldisseri\Irefn{org15}\And
F.~Baltasar Dos Santos Pedrosa\Irefn{org36}\And
R.C.~Baral\Irefn{org60}\And
A.M.~Barbano\Irefn{org110}\And
R.~Barbera\Irefn{org29}\And
F.~Barile\Irefn{org33}\And
G.G.~Barnaf\"{o}ldi\Irefn{org134}\And
L.S.~Barnby\Irefn{org101}\And
V.~Barret\Irefn{org69}\And
P.~Bartalini\Irefn{org7}\And
J.~Bartke\Irefn{org115}\And
E.~Bartsch\Irefn{org52}\And
M.~Basile\Irefn{org28}\And
N.~Bastid\Irefn{org69}\And
S.~Basu\Irefn{org130}\And
B.~Bathen\Irefn{org53}\And
G.~Batigne\Irefn{org112}\And
A.~Batista Camejo\Irefn{org69}\And
B.~Batyunya\Irefn{org65}\And
P.C.~Batzing\Irefn{org22}\And
I.G.~Bearden\Irefn{org79}\And
H.~Beck\Irefn{org52}\And
C.~Bedda\Irefn{org110}\And
N.K.~Behera\Irefn{org48}\textsuperscript{,}\Irefn{org47}\And
I.~Belikov\Irefn{org54}\And
F.~Bellini\Irefn{org28}\And
H.~Bello Martinez\Irefn{org2}\And
R.~Bellwied\Irefn{org120}\And
R.~Belmont\Irefn{org133}\And
E.~Belmont-Moreno\Irefn{org63}\And
V.~Belyaev\Irefn{org75}\And
G.~Bencedi\Irefn{org134}\And
S.~Beole\Irefn{org27}\And
I.~Berceanu\Irefn{org77}\And
A.~Bercuci\Irefn{org77}\And
Y.~Berdnikov\Irefn{org84}\And
D.~Berenyi\Irefn{org134}\And
R.A.~Bertens\Irefn{org56}\And
D.~Berzano\Irefn{org36}\textsuperscript{,}\Irefn{org27}\And
L.~Betev\Irefn{org36}\And
A.~Bhasin\Irefn{org89}\And
I.R.~Bhat\Irefn{org89}\And
A.K.~Bhati\Irefn{org86}\And
B.~Bhattacharjee\Irefn{org44}\And
J.~Bhom\Irefn{org126}\And
L.~Bianchi\Irefn{org27}\textsuperscript{,}\Irefn{org120}\And
N.~Bianchi\Irefn{org71}\And
C.~Bianchin\Irefn{org133}\textsuperscript{,}\Irefn{org56}\And
J.~Biel\v{c}\'{\i}k\Irefn{org39}\And
J.~Biel\v{c}\'{\i}kov\'{a}\Irefn{org82}\And
A.~Bilandzic\Irefn{org79}\And
S.~Biswas\Irefn{org78}\And
S.~Bjelogrlic\Irefn{org56}\And
F.~Blanco\Irefn{org10}\And
D.~Blau\Irefn{org99}\And
C.~Blume\Irefn{org52}\And
F.~Bock\Irefn{org73}\textsuperscript{,}\Irefn{org92}\And
A.~Bogdanov\Irefn{org75}\And
H.~B{\o}ggild\Irefn{org79}\And
L.~Boldizs\'{a}r\Irefn{org134}\And
M.~Bombara\Irefn{org40}\And
J.~Book\Irefn{org52}\And
H.~Borel\Irefn{org15}\And
A.~Borissov\Irefn{org95}\And
M.~Borri\Irefn{org81}\And
F.~Boss\'u\Irefn{org64}\And
M.~Botje\Irefn{org80}\And
E.~Botta\Irefn{org27}\And
S.~B\"{o}ttger\Irefn{org51}\And
P.~Braun-Munzinger\Irefn{org96}\And
M.~Bregant\Irefn{org118}\And
T.~Breitner\Irefn{org51}\And
T.A.~Broker\Irefn{org52}\And
T.A.~Browning\Irefn{org94}\And
M.~Broz\Irefn{org39}\And
E.J.~Brucken\Irefn{org45}\And
E.~Bruna\Irefn{org110}\And
G.E.~Bruno\Irefn{org33}\And
D.~Budnikov\Irefn{org98}\And
H.~Buesching\Irefn{org52}\And
S.~Bufalino\Irefn{org110}\textsuperscript{,}\Irefn{org36}\And
P.~Buncic\Irefn{org36}\And
O.~Busch\Irefn{org92}\textsuperscript{,}\Irefn{org126}\And
Z.~Buthelezi\Irefn{org64}\And
J.T.~Buxton\Irefn{org20}\And
D.~Caffarri\Irefn{org36}\textsuperscript{,}\Irefn{org30}\And
X.~Cai\Irefn{org7}\And
H.~Caines\Irefn{org135}\And
L.~Calero Diaz\Irefn{org71}\And
A.~Caliva\Irefn{org56}\And
E.~Calvo Villar\Irefn{org102}\And
P.~Camerini\Irefn{org26}\And
F.~Carena\Irefn{org36}\And
W.~Carena\Irefn{org36}\And
J.~Castillo Castellanos\Irefn{org15}\And
A.J.~Castro\Irefn{org123}\And
E.A.R.~Casula\Irefn{org25}\And
C.~Cavicchioli\Irefn{org36}\And
C.~Ceballos Sanchez\Irefn{org9}\And
J.~Cepila\Irefn{org39}\And
P.~Cerello\Irefn{org110}\And
B.~Chang\Irefn{org121}\And
S.~Chapeland\Irefn{org36}\And
M.~Chartier\Irefn{org122}\And
J.L.~Charvet\Irefn{org15}\And
S.~Chattopadhyay\Irefn{org130}\And
S.~Chattopadhyay\Irefn{org100}\And
V.~Chelnokov\Irefn{org3}\And
M.~Cherney\Irefn{org85}\And
C.~Cheshkov\Irefn{org128}\And
B.~Cheynis\Irefn{org128}\And
V.~Chibante Barroso\Irefn{org36}\And
D.D.~Chinellato\Irefn{org119}\And
P.~Chochula\Irefn{org36}\And
K.~Choi\Irefn{org95}\And
M.~Chojnacki\Irefn{org79}\And
S.~Choudhury\Irefn{org130}\And
P.~Christakoglou\Irefn{org80}\And
C.H.~Christensen\Irefn{org79}\And
P.~Christiansen\Irefn{org34}\And
T.~Chujo\Irefn{org126}\And
S.U.~Chung\Irefn{org95}\And
Z.~Chunhui\Irefn{org56}\And
C.~Cicalo\Irefn{org105}\And
L.~Cifarelli\Irefn{org12}\textsuperscript{,}\Irefn{org28}\And
F.~Cindolo\Irefn{org104}\And
J.~Cleymans\Irefn{org88}\And
F.~Colamaria\Irefn{org33}\And
D.~Colella\Irefn{org33}\And
A.~Collu\Irefn{org25}\And
M.~Colocci\Irefn{org28}\And
G.~Conesa Balbastre\Irefn{org70}\And
Z.~Conesa del Valle\Irefn{org50}\And
M.E.~Connors\Irefn{org135}\And
J.G.~Contreras\Irefn{org39}\textsuperscript{,}\Irefn{org11}\And
T.M.~Cormier\Irefn{org83}\And
Y.~Corrales Morales\Irefn{org27}\And
I.~Cort\'{e}s Maldonado\Irefn{org2}\And
P.~Cortese\Irefn{org32}\And
M.R.~Cosentino\Irefn{org118}\And
F.~Costa\Irefn{org36}\And
P.~Crochet\Irefn{org69}\And
R.~Cruz Albino\Irefn{org11}\And
E.~Cuautle\Irefn{org62}\And
L.~Cunqueiro\Irefn{org36}\And
T.~Dahms\Irefn{org91}\And
A.~Dainese\Irefn{org107}\And
A.~Danu\Irefn{org61}\And
D.~Das\Irefn{org100}\And
I.~Das\Irefn{org100}\textsuperscript{,}\Irefn{org50}\And
S.~Das\Irefn{org4}\And
A.~Dash\Irefn{org119}\And
S.~Dash\Irefn{org47}\And
S.~De\Irefn{org118}\And
A.~De Caro\Irefn{org31}\textsuperscript{,}\Irefn{org12}\And
G.~de Cataldo\Irefn{org103}\And
J.~de Cuveland\Irefn{org42}\And
A.~De Falco\Irefn{org25}\And
D.~De Gruttola\Irefn{org12}\textsuperscript{,}\Irefn{org31}\And
N.~De Marco\Irefn{org110}\And
S.~De Pasquale\Irefn{org31}\And
A.~Deisting\Irefn{org96}\textsuperscript{,}\Irefn{org92}\And
A.~Deloff\Irefn{org76}\And
E.~D\'{e}nes\Irefn{org134}\And
G.~D'Erasmo\Irefn{org33}\And
D.~Di Bari\Irefn{org33}\And
A.~Di Mauro\Irefn{org36}\And
P.~Di Nezza\Irefn{org71}\And
M.A.~Diaz Corchero\Irefn{org10}\And
T.~Dietel\Irefn{org88}\And
P.~Dillenseger\Irefn{org52}\And
R.~Divi\`{a}\Irefn{org36}\And
{\O}.~Djuvsland\Irefn{org18}\And
A.~Dobrin\Irefn{org56}\textsuperscript{,}\Irefn{org80}\And
T.~Dobrowolski\Irefn{org76}\Aref{0}\And
D.~Domenicis Gimenez\Irefn{org118}\And
B.~D\"{o}nigus\Irefn{org52}\And
O.~Dordic\Irefn{org22}\And
A.K.~Dubey\Irefn{org130}\And
A.~Dubla\Irefn{org56}\And
L.~Ducroux\Irefn{org128}\And
P.~Dupieux\Irefn{org69}\And
R.J.~Ehlers\Irefn{org135}\And
D.~Elia\Irefn{org103}\And
H.~Engel\Irefn{org51}\And
B.~Erazmus\Irefn{org112}\textsuperscript{,}\Irefn{org36}\And
F.~Erhardt\Irefn{org127}\And
D.~Eschweiler\Irefn{org42}\And
B.~Espagnon\Irefn{org50}\And
M.~Estienne\Irefn{org112}\And
S.~Esumi\Irefn{org126}\And
J.~Eum\Irefn{org95}\And
D.~Evans\Irefn{org101}\And
S.~Evdokimov\Irefn{org111}\And
G.~Eyyubova\Irefn{org39}\And
L.~Fabbietti\Irefn{org91}\And
D.~Fabris\Irefn{org107}\And
J.~Faivre\Irefn{org70}\And
A.~Fantoni\Irefn{org71}\And
M.~Fasel\Irefn{org73}\And
L.~Feldkamp\Irefn{org53}\And
D.~Felea\Irefn{org61}\And
A.~Feliciello\Irefn{org110}\And
G.~Feofilov\Irefn{org129}\And
J.~Ferencei\Irefn{org82}\And
A.~Fern\'{a}ndez T\'{e}llez\Irefn{org2}\And
E.G.~Ferreiro\Irefn{org17}\And
A.~Ferretti\Irefn{org27}\And
A.~Festanti\Irefn{org30}\And
J.~Figiel\Irefn{org115}\And
M.A.S.~Figueredo\Irefn{org122}\And
S.~Filchagin\Irefn{org98}\And
D.~Finogeev\Irefn{org55}\And
F.M.~Fionda\Irefn{org103}\And
E.M.~Fiore\Irefn{org33}\And
M.G.~Fleck\Irefn{org92}\And
M.~Floris\Irefn{org36}\And
S.~Foertsch\Irefn{org64}\And
P.~Foka\Irefn{org96}\And
S.~Fokin\Irefn{org99}\And
E.~Fragiacomo\Irefn{org109}\And
A.~Francescon\Irefn{org36}\textsuperscript{,}\Irefn{org30}\And
U.~Frankenfeld\Irefn{org96}\And
U.~Fuchs\Irefn{org36}\And
C.~Furget\Irefn{org70}\And
A.~Furs\Irefn{org55}\And
M.~Fusco Girard\Irefn{org31}\And
J.J.~Gaardh{\o}je\Irefn{org79}\And
M.~Gagliardi\Irefn{org27}\And
A.M.~Gago\Irefn{org102}\And
M.~Gallio\Irefn{org27}\And
D.R.~Gangadharan\Irefn{org73}\And
P.~Ganoti\Irefn{org87}\And
C.~Gao\Irefn{org7}\And
C.~Garabatos\Irefn{org96}\And
E.~Garcia-Solis\Irefn{org13}\And
C.~Gargiulo\Irefn{org36}\And
P.~Gasik\Irefn{org91}\And
M.~Germain\Irefn{org112}\And
A.~Gheata\Irefn{org36}\And
M.~Gheata\Irefn{org61}\textsuperscript{,}\Irefn{org36}\And
P.~Ghosh\Irefn{org130}\And
S.K.~Ghosh\Irefn{org4}\And
P.~Gianotti\Irefn{org71}\And
P.~Giubellino\Irefn{org36}\And
P.~Giubilato\Irefn{org30}\And
E.~Gladysz-Dziadus\Irefn{org115}\And
P.~Gl\"{a}ssel\Irefn{org92}\And
A.~Gomez Ramirez\Irefn{org51}\And
P.~Gonz\'{a}lez-Zamora\Irefn{org10}\And
S.~Gorbunov\Irefn{org42}\And
L.~G\"{o}rlich\Irefn{org115}\And
S.~Gotovac\Irefn{org114}\And
V.~Grabski\Irefn{org63}\And
L.K.~Graczykowski\Irefn{org132}\And
A.~Grelli\Irefn{org56}\And
A.~Grigoras\Irefn{org36}\And
C.~Grigoras\Irefn{org36}\And
V.~Grigoriev\Irefn{org75}\And
A.~Grigoryan\Irefn{org1}\And
S.~Grigoryan\Irefn{org65}\And
B.~Grinyov\Irefn{org3}\And
N.~Grion\Irefn{org109}\And
J.F.~Grosse-Oetringhaus\Irefn{org36}\And
J.-Y.~Grossiord\Irefn{org128}\And
R.~Grosso\Irefn{org36}\And
F.~Guber\Irefn{org55}\And
R.~Guernane\Irefn{org70}\And
B.~Guerzoni\Irefn{org28}\And
K.~Gulbrandsen\Irefn{org79}\And
H.~Gulkanyan\Irefn{org1}\And
T.~Gunji\Irefn{org125}\And
A.~Gupta\Irefn{org89}\And
R.~Gupta\Irefn{org89}\And
R.~Haake\Irefn{org53}\And
{\O}.~Haaland\Irefn{org18}\And
C.~Hadjidakis\Irefn{org50}\And
M.~Haiduc\Irefn{org61}\And
H.~Hamagaki\Irefn{org125}\And
G.~Hamar\Irefn{org134}\And
L.D.~Hanratty\Irefn{org101}\And
A.~Hansen\Irefn{org79}\And
J.W.~Harris\Irefn{org135}\And
H.~Hartmann\Irefn{org42}\And
A.~Harton\Irefn{org13}\And
D.~Hatzifotiadou\Irefn{org104}\And
S.~Hayashi\Irefn{org125}\And
S.T.~Heckel\Irefn{org52}\And
M.~Heide\Irefn{org53}\And
H.~Helstrup\Irefn{org37}\And
A.~Herghelegiu\Irefn{org77}\And
G.~Herrera Corral\Irefn{org11}\And
B.A.~Hess\Irefn{org35}\And
K.F.~Hetland\Irefn{org37}\And
T.E.~Hilden\Irefn{org45}\And
H.~Hillemanns\Irefn{org36}\And
B.~Hippolyte\Irefn{org54}\And
P.~Hristov\Irefn{org36}\And
M.~Huang\Irefn{org18}\And
T.J.~Humanic\Irefn{org20}\And
N.~Hussain\Irefn{org44}\And
T.~Hussain\Irefn{org19}\And
D.~Hutter\Irefn{org42}\And
D.S.~Hwang\Irefn{org21}\And
R.~Ilkaev\Irefn{org98}\And
I.~Ilkiv\Irefn{org76}\And
M.~Inaba\Irefn{org126}\And
C.~Ionita\Irefn{org36}\And
M.~Ippolitov\Irefn{org75}\textsuperscript{,}\Irefn{org99}\And
M.~Irfan\Irefn{org19}\And
M.~Ivanov\Irefn{org96}\And
V.~Ivanov\Irefn{org84}\And
V.~Izucheev\Irefn{org111}\And
P.M.~Jacobs\Irefn{org73}\And
C.~Jahnke\Irefn{org118}\And
H.J.~Jang\Irefn{org67}\And
M.A.~Janik\Irefn{org132}\And
P.H.S.Y.~Jayarathna\Irefn{org120}\And
C.~Jena\Irefn{org30}\And
S.~Jena\Irefn{org120}\And
R.T.~Jimenez Bustamante\Irefn{org62}\And
P.G.~Jones\Irefn{org101}\And
H.~Jung\Irefn{org43}\And
A.~Jusko\Irefn{org101}\And
P.~Kalinak\Irefn{org58}\And
A.~Kalweit\Irefn{org36}\And
J.~Kamin\Irefn{org52}\And
J.H.~Kang\Irefn{org136}\And
V.~Kaplin\Irefn{org75}\And
S.~Kar\Irefn{org130}\And
A.~Karasu Uysal\Irefn{org68}\And
O.~Karavichev\Irefn{org55}\And
T.~Karavicheva\Irefn{org55}\And
E.~Karpechev\Irefn{org55}\And
U.~Kebschull\Irefn{org51}\And
R.~Keidel\Irefn{org137}\And
D.L.D.~Keijdener\Irefn{org56}\And
M.~Keil\Irefn{org36}\And
K.H.~Khan\Irefn{org16}\And
M.M.~Khan\Irefn{org19}\And
P.~Khan\Irefn{org100}\And
S.A.~Khan\Irefn{org130}\And
A.~Khanzadeev\Irefn{org84}\And
Y.~Kharlov\Irefn{org111}\And
B.~Kileng\Irefn{org37}\And
B.~Kim\Irefn{org136}\And
D.W.~Kim\Irefn{org43}\textsuperscript{,}\Irefn{org67}\And
D.J.~Kim\Irefn{org121}\And
H.~Kim\Irefn{org136}\And
J.S.~Kim\Irefn{org43}\And
M.~Kim\Irefn{org43}\And
M.~Kim\Irefn{org136}\And
S.~Kim\Irefn{org21}\And
T.~Kim\Irefn{org136}\And
S.~Kirsch\Irefn{org42}\And
I.~Kisel\Irefn{org42}\And
S.~Kiselev\Irefn{org57}\And
A.~Kisiel\Irefn{org132}\And
G.~Kiss\Irefn{org134}\And
J.L.~Klay\Irefn{org6}\And
C.~Klein\Irefn{org52}\And
J.~Klein\Irefn{org92}\And
C.~Klein-B\"{o}sing\Irefn{org53}\And
A.~Kluge\Irefn{org36}\And
M.L.~Knichel\Irefn{org92}\And
A.G.~Knospe\Irefn{org116}\And
T.~Kobayashi\Irefn{org126}\And
C.~Kobdaj\Irefn{org113}\And
M.~Kofarago\Irefn{org36}\And
M.K.~K\"{o}hler\Irefn{org96}\And
T.~Kollegger\Irefn{org42}\textsuperscript{,}\Irefn{org96}\And
A.~Kolojvari\Irefn{org129}\And
V.~Kondratiev\Irefn{org129}\And
N.~Kondratyeva\Irefn{org75}\And
E.~Kondratyuk\Irefn{org111}\And
A.~Konevskikh\Irefn{org55}\And
C.~Kouzinopoulos\Irefn{org36}\And
O.~Kovalenko\Irefn{org76}\And
V.~Kovalenko\Irefn{org129}\And
M.~Kowalski\Irefn{org36}\textsuperscript{,}\Irefn{org115}\And
S.~Kox\Irefn{org70}\And
G.~Koyithatta Meethaleveedu\Irefn{org47}\And
J.~Kral\Irefn{org121}\And
I.~Kr\'{a}lik\Irefn{org58}\And
A.~Krav\v{c}\'{a}kov\'{a}\Irefn{org40}\And
M.~Krelina\Irefn{org39}\And
M.~Kretz\Irefn{org42}\And
M.~Krivda\Irefn{org101}\textsuperscript{,}\Irefn{org58}\And
F.~Krizek\Irefn{org82}\And
E.~Kryshen\Irefn{org36}\And
M.~Krzewicki\Irefn{org96}\textsuperscript{,}\Irefn{org42}\And
A.M.~Kubera\Irefn{org20}\And
V.~Ku\v{c}era\Irefn{org82}\And
T.~Kugathasan\Irefn{org36}\And
C.~Kuhn\Irefn{org54}\And
P.G.~Kuijer\Irefn{org80}\And
I.~Kulakov\Irefn{org42}\And
J.~Kumar\Irefn{org47}\And
L.~Kumar\Irefn{org78}\textsuperscript{,}\Irefn{org86}\And
P.~Kurashvili\Irefn{org76}\And
A.~Kurepin\Irefn{org55}\And
A.B.~Kurepin\Irefn{org55}\And
A.~Kuryakin\Irefn{org98}\And
S.~Kushpil\Irefn{org82}\And
M.J.~Kweon\Irefn{org49}\And
Y.~Kwon\Irefn{org136}\And
S.L.~La Pointe\Irefn{org110}\And
P.~La Rocca\Irefn{org29}\And
C.~Lagana Fernandes\Irefn{org118}\And
I.~Lakomov\Irefn{org36}\textsuperscript{,}\Irefn{org50}\And
R.~Langoy\Irefn{org41}\And
C.~Lara\Irefn{org51}\And
A.~Lardeux\Irefn{org15}\And
A.~Lattuca\Irefn{org27}\And
E.~Laudi\Irefn{org36}\And
R.~Lea\Irefn{org26}\And
L.~Leardini\Irefn{org92}\And
G.R.~Lee\Irefn{org101}\And
S.~Lee\Irefn{org136}\And
I.~Legrand\Irefn{org36}\And
R.C.~Lemmon\Irefn{org81}\And
V.~Lenti\Irefn{org103}\And
E.~Leogrande\Irefn{org56}\And
I.~Le\'{o}n Monz\'{o}n\Irefn{org117}\And
M.~Leoncino\Irefn{org27}\And
P.~L\'{e}vai\Irefn{org134}\And
S.~Li\Irefn{org7}\textsuperscript{,}\Irefn{org69}\And
X.~Li\Irefn{org14}\And
J.~Lien\Irefn{org41}\And
R.~Lietava\Irefn{org101}\And
S.~Lindal\Irefn{org22}\And
V.~Lindenstruth\Irefn{org42}\And
C.~Lippmann\Irefn{org96}\And
M.A.~Lisa\Irefn{org20}\And
H.M.~Ljunggren\Irefn{org34}\And
D.F.~Lodato\Irefn{org56}\And
P.I.~Loenne\Irefn{org18}\And
V.R.~Loggins\Irefn{org133}\And
V.~Loginov\Irefn{org75}\And
C.~Loizides\Irefn{org73}\And
X.~Lopez\Irefn{org69}\And
E.~L\'{o}pez Torres\Irefn{org9}\And
A.~Lowe\Irefn{org134}\And
P.~Luettig\Irefn{org52}\And
M.~Lunardon\Irefn{org30}\And
G.~Luparello\Irefn{org26}\textsuperscript{,}\Irefn{org56}\And
P.H.F.N.D.~Luz\Irefn{org118}\And
A.~Maevskaya\Irefn{org55}\And
M.~Mager\Irefn{org36}\And
S.~Mahajan\Irefn{org89}\And
S.M.~Mahmood\Irefn{org22}\And
A.~Maire\Irefn{org54}\And
R.D.~Majka\Irefn{org135}\And
M.~Malaev\Irefn{org84}\And
I.~Maldonado Cervantes\Irefn{org62}\And
L.~Malinina\Irefn{org65}\And
D.~Mal'Kevich\Irefn{org57}\And
P.~Malzacher\Irefn{org96}\And
A.~Mamonov\Irefn{org98}\And
L.~Manceau\Irefn{org110}\And
V.~Manko\Irefn{org99}\And
F.~Manso\Irefn{org69}\And
V.~Manzari\Irefn{org103}\textsuperscript{,}\Irefn{org36}\And
M.~Marchisone\Irefn{org27}\And
J.~Mare\v{s}\Irefn{org59}\And
G.V.~Margagliotti\Irefn{org26}\And
A.~Margotti\Irefn{org104}\And
J.~Margutti\Irefn{org56}\And
A.~Mar\'{\i}n\Irefn{org96}\And
C.~Markert\Irefn{org116}\And
M.~Marquard\Irefn{org52}\And
N.A.~Martin\Irefn{org96}\And
J.~Martin Blanco\Irefn{org112}\And
P.~Martinengo\Irefn{org36}\And
M.I.~Mart\'{\i}nez\Irefn{org2}\And
G.~Mart\'{\i}nez Garc\'{\i}a\Irefn{org112}\And
M.~Martinez Pedreira\Irefn{org36}\And
Y.~Martynov\Irefn{org3}\And
A.~Mas\Irefn{org118}\And
S.~Masciocchi\Irefn{org96}\And
M.~Masera\Irefn{org27}\And
A.~Masoni\Irefn{org105}\And
L.~Massacrier\Irefn{org112}\And
A.~Mastroserio\Irefn{org33}\And
H.~Masui\Irefn{org126}\And
A.~Matyja\Irefn{org115}\And
C.~Mayer\Irefn{org115}\And
J.~Mazer\Irefn{org123}\And
M.A.~Mazzoni\Irefn{org108}\And
D.~Mcdonald\Irefn{org120}\And
F.~Meddi\Irefn{org24}\And
A.~Menchaca-Rocha\Irefn{org63}\And
E.~Meninno\Irefn{org31}\And
J.~Mercado P\'erez\Irefn{org92}\And
M.~Meres\Irefn{org38}\And
Y.~Miake\Irefn{org126}\And
M.M.~Mieskolainen\Irefn{org45}\And
K.~Mikhaylov\Irefn{org57}\textsuperscript{,}\Irefn{org65}\And
L.~Milano\Irefn{org36}\And
J.~Milosevic\Irefn{org22}\textsuperscript{,}\Irefn{org131}\And
L.M.~Minervini\Irefn{org103}\textsuperscript{,}\Irefn{org23}\And
A.~Mischke\Irefn{org56}\And
A.N.~Mishra\Irefn{org48}\And
D.~Mi\'{s}kowiec\Irefn{org96}\And
J.~Mitra\Irefn{org130}\And
C.M.~Mitu\Irefn{org61}\And
N.~Mohammadi\Irefn{org56}\And
B.~Mohanty\Irefn{org130}\textsuperscript{,}\Irefn{org78}\And
L.~Molnar\Irefn{org54}\And
L.~Monta\~{n}o Zetina\Irefn{org11}\And
E.~Montes\Irefn{org10}\And
M.~Morando\Irefn{org30}\And
D.A.~Moreira De Godoy\Irefn{org112}\And
S.~Moretto\Irefn{org30}\And
A.~Morreale\Irefn{org112}\And
A.~Morsch\Irefn{org36}\And
V.~Muccifora\Irefn{org71}\And
E.~Mudnic\Irefn{org114}\And
D.~M{\"u}hlheim\Irefn{org53}\And
S.~Muhuri\Irefn{org130}\And
M.~Mukherjee\Irefn{org130}\And
H.~M\"{u}ller\Irefn{org36}\And
J.D.~Mulligan\Irefn{org135}\And
M.G.~Munhoz\Irefn{org118}\And
S.~Murray\Irefn{org64}\And
L.~Musa\Irefn{org36}\And
J.~Musinsky\Irefn{org58}\And
B.K.~Nandi\Irefn{org47}\And
R.~Nania\Irefn{org104}\And
E.~Nappi\Irefn{org103}\And
M.U.~Naru\Irefn{org16}\And
C.~Nattrass\Irefn{org123}\And
K.~Nayak\Irefn{org78}\And
T.K.~Nayak\Irefn{org130}\And
S.~Nazarenko\Irefn{org98}\And
A.~Nedosekin\Irefn{org57}\And
L.~Nellen\Irefn{org62}\And
F.~Ng\Irefn{org120}\And
M.~Nicassio\Irefn{org96}\And
M.~Niculescu\Irefn{org61}\textsuperscript{,}\Irefn{org36}\And
J.~Niedziela\Irefn{org36}\And
B.S.~Nielsen\Irefn{org79}\And
S.~Nikolaev\Irefn{org99}\And
S.~Nikulin\Irefn{org99}\And
V.~Nikulin\Irefn{org84}\And
F.~Noferini\Irefn{org104}\textsuperscript{,}\Irefn{org12}\And
P.~Nomokonov\Irefn{org65}\And
G.~Nooren\Irefn{org56}\And
J.~Norman\Irefn{org122}\And
A.~Nyanin\Irefn{org99}\And
J.~Nystrand\Irefn{org18}\And
H.~Oeschler\Irefn{org92}\And
S.~Oh\Irefn{org135}\And
S.K.~Oh\Irefn{org66}\And
A.~Ohlson\Irefn{org36}\And
A.~Okatan\Irefn{org68}\And
T.~Okubo\Irefn{org46}\And
L.~Olah\Irefn{org134}\And
J.~Oleniacz\Irefn{org132}\And
A.C.~Oliveira Da Silva\Irefn{org118}\And
M.H.~Oliver\Irefn{org135}\And
J.~Onderwaater\Irefn{org96}\And
C.~Oppedisano\Irefn{org110}\And
A.~Ortiz Velasquez\Irefn{org62}\And
A.~Oskarsson\Irefn{org34}\And
J.~Otwinowski\Irefn{org96}\textsuperscript{,}\Irefn{org115}\And
K.~Oyama\Irefn{org92}\And
M.~Ozdemir\Irefn{org52}\And
Y.~Pachmayer\Irefn{org92}\And
P.~Pagano\Irefn{org31}\And
G.~Pai\'{c}\Irefn{org62}\And
C.~Pajares\Irefn{org17}\And
S.K.~Pal\Irefn{org130}\And
J.~Pan\Irefn{org133}\And
A.K.~Pandey\Irefn{org47}\And
D.~Pant\Irefn{org47}\And
V.~Papikyan\Irefn{org1}\And
G.S.~Pappalardo\Irefn{org106}\And
P.~Pareek\Irefn{org48}\And
W.J.~Park\Irefn{org96}\And
S.~Parmar\Irefn{org86}\And
A.~Passfeld\Irefn{org53}\And
V.~Paticchio\Irefn{org103}\And
B.~Paul\Irefn{org100}\And
T.~Peitzmann\Irefn{org56}\And
H.~Pereira Da Costa\Irefn{org15}\And
E.~Pereira De Oliveira Filho\Irefn{org118}\And
D.~Peresunko\Irefn{org75}\textsuperscript{,}\Irefn{org99}\And
C.E.~P\'erez Lara\Irefn{org80}\And
V.~Peskov\Irefn{org52}\And
Y.~Pestov\Irefn{org5}\And
V.~Petr\'{a}\v{c}ek\Irefn{org39}\And
V.~Petrov\Irefn{org111}\And
M.~Petrovici\Irefn{org77}\And
C.~Petta\Irefn{org29}\And
S.~Piano\Irefn{org109}\And
M.~Pikna\Irefn{org38}\And
P.~Pillot\Irefn{org112}\And
O.~Pinazza\Irefn{org104}\textsuperscript{,}\Irefn{org36}\And
L.~Pinsky\Irefn{org120}\And
D.B.~Piyarathna\Irefn{org120}\And
M.~P\l osko\'{n}\Irefn{org73}\And
M.~Planinic\Irefn{org127}\And
J.~Pluta\Irefn{org132}\And
S.~Pochybova\Irefn{org134}\And
P.L.M.~Podesta-Lerma\Irefn{org117}\And
M.G.~Poghosyan\Irefn{org85}\And
B.~Polichtchouk\Irefn{org111}\And
N.~Poljak\Irefn{org127}\And
W.~Poonsawat\Irefn{org113}\And
A.~Pop\Irefn{org77}\And
S.~Porteboeuf-Houssais\Irefn{org69}\And
J.~Porter\Irefn{org73}\And
J.~Pospisil\Irefn{org82}\And
S.K.~Prasad\Irefn{org4}\And
R.~Preghenella\Irefn{org36}\textsuperscript{,}\Irefn{org104}\And
F.~Prino\Irefn{org110}\And
C.A.~Pruneau\Irefn{org133}\And
I.~Pshenichnov\Irefn{org55}\And
M.~Puccio\Irefn{org110}\And
G.~Puddu\Irefn{org25}\And
P.~Pujahari\Irefn{org133}\And
V.~Punin\Irefn{org98}\And
J.~Putschke\Irefn{org133}\And
H.~Qvigstad\Irefn{org22}\And
A.~Rachevski\Irefn{org109}\And
S.~Raha\Irefn{org4}\And
S.~Rajput\Irefn{org89}\And
J.~Rak\Irefn{org121}\And
A.~Rakotozafindrabe\Irefn{org15}\And
L.~Ramello\Irefn{org32}\And
R.~Raniwala\Irefn{org90}\And
S.~Raniwala\Irefn{org90}\And
S.S.~R\"{a}s\"{a}nen\Irefn{org45}\And
B.T.~Rascanu\Irefn{org52}\And
D.~Rathee\Irefn{org86}\And
K.F.~Read\Irefn{org123}\And
J.S.~Real\Irefn{org70}\And
K.~Redlich\Irefn{org76}\And
R.J.~Reed\Irefn{org133}\And
A.~Rehman\Irefn{org18}\And
P.~Reichelt\Irefn{org52}\And
M.~Reicher\Irefn{org56}\And
F.~Reidt\Irefn{org92}\textsuperscript{,}\Irefn{org36}\And
X.~Ren\Irefn{org7}\And
R.~Renfordt\Irefn{org52}\And
A.R.~Reolon\Irefn{org71}\And
A.~Reshetin\Irefn{org55}\And
F.~Rettig\Irefn{org42}\And
J.-P.~Revol\Irefn{org12}\And
K.~Reygers\Irefn{org92}\And
V.~Riabov\Irefn{org84}\And
R.A.~Ricci\Irefn{org72}\And
T.~Richert\Irefn{org34}\And
M.~Richter\Irefn{org22}\And
P.~Riedler\Irefn{org36}\And
W.~Riegler\Irefn{org36}\And
F.~Riggi\Irefn{org29}\And
C.~Ristea\Irefn{org61}\And
A.~Rivetti\Irefn{org110}\And
E.~Rocco\Irefn{org56}\And
M.~Rodr\'{i}guez Cahuantzi\Irefn{org11}\textsuperscript{,}\Irefn{org2}\And
A.~Rodriguez Manso\Irefn{org80}\And
K.~R{\o}ed\Irefn{org22}\And
E.~Rogochaya\Irefn{org65}\And
D.~Rohr\Irefn{org42}\And
D.~R\"ohrich\Irefn{org18}\And
R.~Romita\Irefn{org122}\And
F.~Ronchetti\Irefn{org71}\And
L.~Ronflette\Irefn{org112}\And
P.~Rosnet\Irefn{org69}\And
A.~Rossi\Irefn{org36}\And
F.~Roukoutakis\Irefn{org87}\And
A.~Roy\Irefn{org48}\And
C.~Roy\Irefn{org54}\And
P.~Roy\Irefn{org100}\And
A.J.~Rubio Montero\Irefn{org10}\And
R.~Rui\Irefn{org26}\And
R.~Russo\Irefn{org27}\And
E.~Ryabinkin\Irefn{org99}\And
Y.~Ryabov\Irefn{org84}\And
A.~Rybicki\Irefn{org115}\And
S.~Sadovsky\Irefn{org111}\And
K.~\v{S}afa\v{r}\'{\i}k\Irefn{org36}\And
B.~Sahlmuller\Irefn{org52}\And
P.~Sahoo\Irefn{org48}\And
R.~Sahoo\Irefn{org48}\And
S.~Sahoo\Irefn{org60}\And
P.K.~Sahu\Irefn{org60}\And
J.~Saini\Irefn{org130}\And
S.~Sakai\Irefn{org71}\And
M.A.~Saleh\Irefn{org133}\And
C.A.~Salgado\Irefn{org17}\And
J.~Salzwedel\Irefn{org20}\And
S.~Sambyal\Irefn{org89}\And
V.~Samsonov\Irefn{org84}\And
X.~Sanchez Castro\Irefn{org54}\And
L.~\v{S}\'{a}ndor\Irefn{org58}\And
A.~Sandoval\Irefn{org63}\And
M.~Sano\Irefn{org126}\And
G.~Santagati\Irefn{org29}\And
D.~Sarkar\Irefn{org130}\And
E.~Scapparone\Irefn{org104}\And
F.~Scarlassara\Irefn{org30}\And
R.P.~Scharenberg\Irefn{org94}\And
C.~Schiaua\Irefn{org77}\And
R.~Schicker\Irefn{org92}\And
C.~Schmidt\Irefn{org96}\And
H.R.~Schmidt\Irefn{org35}\And
S.~Schuchmann\Irefn{org52}\And
J.~Schukraft\Irefn{org36}\And
M.~Schulc\Irefn{org39}\And
T.~Schuster\Irefn{org135}\And
Y.~Schutz\Irefn{org112}\textsuperscript{,}\Irefn{org36}\And
K.~Schwarz\Irefn{org96}\And
K.~Schweda\Irefn{org96}\And
G.~Scioli\Irefn{org28}\And
E.~Scomparin\Irefn{org110}\And
R.~Scott\Irefn{org123}\And
K.S.~Seeder\Irefn{org118}\And
J.E.~Seger\Irefn{org85}\And
Y.~Sekiguchi\Irefn{org125}\And
I.~Selyuzhenkov\Irefn{org96}\And
K.~Senosi\Irefn{org64}\And
J.~Seo\Irefn{org66}\textsuperscript{,}\Irefn{org95}\And
E.~Serradilla\Irefn{org10}\textsuperscript{,}\Irefn{org63}\And
A.~Sevcenco\Irefn{org61}\And
A.~Shabanov\Irefn{org55}\And
A.~Shabetai\Irefn{org112}\And
O.~Shadura\Irefn{org3}\And
R.~Shahoyan\Irefn{org36}\And
A.~Shangaraev\Irefn{org111}\And
A.~Sharma\Irefn{org89}\And
N.~Sharma\Irefn{org60}\textsuperscript{,}\Irefn{org123}\And
K.~Shigaki\Irefn{org46}\And
K.~Shtejer\Irefn{org9}\textsuperscript{,}\Irefn{org27}\And
Y.~Sibiriak\Irefn{org99}\And
S.~Siddhanta\Irefn{org105}\And
K.M.~Sielewicz\Irefn{org36}\And
T.~Siemiarczuk\Irefn{org76}\And
D.~Silvermyr\Irefn{org83}\textsuperscript{,}\Irefn{org34}\And
C.~Silvestre\Irefn{org70}\And
G.~Simatovic\Irefn{org127}\And
G.~Simonetti\Irefn{org36}\And
R.~Singaraju\Irefn{org130}\And
R.~Singh\Irefn{org78}\And
S.~Singha\Irefn{org78}\textsuperscript{,}\Irefn{org130}\And
V.~Singhal\Irefn{org130}\And
B.C.~Sinha\Irefn{org130}\And
T.~Sinha\Irefn{org100}\And
B.~Sitar\Irefn{org38}\And
M.~Sitta\Irefn{org32}\And
T.B.~Skaali\Irefn{org22}\And
M.~Slupecki\Irefn{org121}\And
N.~Smirnov\Irefn{org135}\And
R.J.M.~Snellings\Irefn{org56}\And
T.W.~Snellman\Irefn{org121}\And
C.~S{\o}gaard\Irefn{org34}\And
R.~Soltz\Irefn{org74}\And
J.~Song\Irefn{org95}\And
M.~Song\Irefn{org136}\And
Z.~Song\Irefn{org7}\And
F.~Soramel\Irefn{org30}\And
S.~Sorensen\Irefn{org123}\And
M.~Spacek\Irefn{org39}\And
E.~Spiriti\Irefn{org71}\And
I.~Sputowska\Irefn{org115}\And
M.~Spyropoulou-Stassinaki\Irefn{org87}\And
B.K.~Srivastava\Irefn{org94}\And
J.~Stachel\Irefn{org92}\And
I.~Stan\Irefn{org61}\And
G.~Stefanek\Irefn{org76}\And
M.~Steinpreis\Irefn{org20}\And
E.~Stenlund\Irefn{org34}\And
G.~Steyn\Irefn{org64}\And
J.H.~Stiller\Irefn{org92}\And
D.~Stocco\Irefn{org112}\And
P.~Strmen\Irefn{org38}\And
A.A.P.~Suaide\Irefn{org118}\And
T.~Sugitate\Irefn{org46}\And
C.~Suire\Irefn{org50}\And
M.~Suleymanov\Irefn{org16}\And
R.~Sultanov\Irefn{org57}\And
M.~\v{S}umbera\Irefn{org82}\And
T.J.M.~Symons\Irefn{org73}\And
A.~Szabo\Irefn{org38}\And
A.~Szanto de Toledo\Irefn{org118}\And
I.~Szarka\Irefn{org38}\And
A.~Szczepankiewicz\Irefn{org36}\And
M.~Szymanski\Irefn{org132}\And
J.~Takahashi\Irefn{org119}\And
N.~Tanaka\Irefn{org126}\And
M.A.~Tangaro\Irefn{org33}\And
J.D.~Tapia Takaki\Aref{idp5842272}\textsuperscript{,}\Irefn{org50}\And
A.~Tarantola Peloni\Irefn{org52}\And
M.~Tariq\Irefn{org19}\And
M.G.~Tarzila\Irefn{org77}\And
A.~Tauro\Irefn{org36}\And
G.~Tejeda Mu\~{n}oz\Irefn{org2}\And
A.~Telesca\Irefn{org36}\And
K.~Terasaki\Irefn{org125}\And
C.~Terrevoli\Irefn{org30}\textsuperscript{,}\Irefn{org25}\And
B.~Teyssier\Irefn{org128}\And
J.~Th\"{a}der\Irefn{org96}\textsuperscript{,}\Irefn{org73}\And
D.~Thomas\Irefn{org116}\And
R.~Tieulent\Irefn{org128}\And
A.R.~Timmins\Irefn{org120}\And
A.~Toia\Irefn{org52}\And
S.~Trogolo\Irefn{org110}\And
V.~Trubnikov\Irefn{org3}\And
W.H.~Trzaska\Irefn{org121}\And
T.~Tsuji\Irefn{org125}\And
A.~Tumkin\Irefn{org98}\And
R.~Turrisi\Irefn{org107}\And
T.S.~Tveter\Irefn{org22}\And
K.~Ullaland\Irefn{org18}\And
A.~Uras\Irefn{org128}\And
G.L.~Usai\Irefn{org25}\And
A.~Utrobicic\Irefn{org127}\And
M.~Vajzer\Irefn{org82}\And
M.~Vala\Irefn{org58}\And
L.~Valencia Palomo\Irefn{org69}\And
S.~Vallero\Irefn{org27}\And
J.~Van Der Maarel\Irefn{org56}\And
J.W.~Van Hoorne\Irefn{org36}\And
M.~van Leeuwen\Irefn{org56}\And
T.~Vanat\Irefn{org82}\And
P.~Vande Vyvre\Irefn{org36}\And
D.~Varga\Irefn{org134}\And
A.~Vargas\Irefn{org2}\And
M.~Vargyas\Irefn{org121}\And
R.~Varma\Irefn{org47}\And
M.~Vasileiou\Irefn{org87}\And
A.~Vasiliev\Irefn{org99}\And
A.~Vauthier\Irefn{org70}\And
V.~Vechernin\Irefn{org129}\And
A.M.~Veen\Irefn{org56}\And
M.~Veldhoen\Irefn{org56}\And
A.~Velure\Irefn{org18}\And
M.~Venaruzzo\Irefn{org72}\And
E.~Vercellin\Irefn{org27}\And
S.~Vergara Lim\'on\Irefn{org2}\And
R.~Vernet\Irefn{org8}\And
M.~Verweij\Irefn{org133}\And
L.~Vickovic\Irefn{org114}\And
G.~Viesti\Irefn{org30}\Aref{0}\And
J.~Viinikainen\Irefn{org121}\And
Z.~Vilakazi\Irefn{org124}\And
O.~Villalobos Baillie\Irefn{org101}\And
A.~Vinogradov\Irefn{org99}\And
L.~Vinogradov\Irefn{org129}\And
Y.~Vinogradov\Irefn{org98}\And
T.~Virgili\Irefn{org31}\And
V.~Vislavicius\Irefn{org34}\And
Y.P.~Viyogi\Irefn{org130}\And
A.~Vodopyanov\Irefn{org65}\And
M.A.~V\"{o}lkl\Irefn{org92}\And
K.~Voloshin\Irefn{org57}\And
S.A.~Voloshin\Irefn{org133}\And
G.~Volpe\Irefn{org36}\textsuperscript{,}\Irefn{org134}\And
B.~von Haller\Irefn{org36}\And
I.~Vorobyev\Irefn{org91}\And
D.~Vranic\Irefn{org96}\textsuperscript{,}\Irefn{org36}\And
J.~Vrl\'{a}kov\'{a}\Irefn{org40}\And
B.~Vulpescu\Irefn{org69}\And
A.~Vyushin\Irefn{org98}\And
B.~Wagner\Irefn{org18}\And
J.~Wagner\Irefn{org96}\And
H.~Wang\Irefn{org56}\And
M.~Wang\Irefn{org7}\textsuperscript{,}\Irefn{org112}\And
Y.~Wang\Irefn{org92}\And
D.~Watanabe\Irefn{org126}\And
M.~Weber\Irefn{org36}\And
S.G.~Weber\Irefn{org96}\And
J.P.~Wessels\Irefn{org53}\And
U.~Westerhoff\Irefn{org53}\And
J.~Wiechula\Irefn{org35}\And
J.~Wikne\Irefn{org22}\And
M.~Wilde\Irefn{org53}\And
G.~Wilk\Irefn{org76}\And
J.~Wilkinson\Irefn{org92}\And
M.C.S.~Williams\Irefn{org104}\And
B.~Windelband\Irefn{org92}\And
M.~Winn\Irefn{org92}\And
C.G.~Yaldo\Irefn{org133}\And
Y.~Yamaguchi\Irefn{org125}\And
H.~Yang\Irefn{org56}\And
P.~Yang\Irefn{org7}\And
S.~Yano\Irefn{org46}\And
Z.~Yin\Irefn{org7}\And
H.~Yokoyama\Irefn{org126}\And
I.-K.~Yoo\Irefn{org95}\And
V.~Yurchenko\Irefn{org3}\And
I.~Yushmanov\Irefn{org99}\And
A.~Zaborowska\Irefn{org132}\And
V.~Zaccolo\Irefn{org79}\And
A.~Zaman\Irefn{org16}\And
C.~Zampolli\Irefn{org104}\And
H.J.C.~Zanoli\Irefn{org118}\And
S.~Zaporozhets\Irefn{org65}\And
A.~Zarochentsev\Irefn{org129}\And
P.~Z\'{a}vada\Irefn{org59}\And
N.~Zaviyalov\Irefn{org98}\And
H.~Zbroszczyk\Irefn{org132}\And
I.S.~Zgura\Irefn{org61}\And
M.~Zhalov\Irefn{org84}\And
H.~Zhang\Irefn{org7}\And
X.~Zhang\Irefn{org73}\And
Y.~Zhang\Irefn{org7}\And
C.~Zhao\Irefn{org22}\And
N.~Zhigareva\Irefn{org57}\And
D.~Zhou\Irefn{org7}\And
Y.~Zhou\Irefn{org56}\And
Z.~Zhou\Irefn{org18}\And
H.~Zhu\Irefn{org7}\And
J.~Zhu\Irefn{org7}\textsuperscript{,}\Irefn{org112}\And
X.~Zhu\Irefn{org7}\And
A.~Zichichi\Irefn{org12}\textsuperscript{,}\Irefn{org28}\And
A.~Zimmermann\Irefn{org92}\And
M.B.~Zimmermann\Irefn{org53}\textsuperscript{,}\Irefn{org36}\And
G.~Zinovjev\Irefn{org3}\And
M.~Zyzak\Irefn{org42}
\renewcommand\labelenumi{\textsuperscript{\theenumi}~}

\section*{Affiliation notes}
\renewcommand\theenumi{\roman{enumi}}
\begin{Authlist}
\item \Adef{0}Deceased
\item \Adef{idp5842272}{Also at: University of Kansas, Lawrence, Kansas, United States}
\end{Authlist}

\section*{Collaboration Institutes}
\renewcommand\theenumi{\arabic{enumi}~}
\begin{Authlist}

\item \Idef{org1}A.I. Alikhanyan National Science Laboratory (Yerevan Physics Institute) Foundation, Yerevan, Armenia
\item \Idef{org2}Benem\'{e}rita Universidad Aut\'{o}noma de Puebla, Puebla, Mexico
\item \Idef{org3}Bogolyubov Institute for Theoretical Physics, Kiev, Ukraine
\item \Idef{org4}Bose Institute, Department of Physics and Centre for Astroparticle Physics and Space Science (CAPSS), Kolkata, India
\item \Idef{org5}Budker Institute for Nuclear Physics, Novosibirsk, Russia
\item \Idef{org6}California Polytechnic State University, San Luis Obispo, California, United States
\item \Idef{org7}Central China Normal University, Wuhan, China
\item \Idef{org8}Centre de Calcul de l'IN2P3, Villeurbanne, France
\item \Idef{org9}Centro de Aplicaciones Tecnol\'{o}gicas y Desarrollo Nuclear (CEADEN), Havana, Cuba
\item \Idef{org10}Centro de Investigaciones Energ\'{e}ticas Medioambientales y Tecnol\'{o}gicas (CIEMAT), Madrid, Spain
\item \Idef{org11}Centro de Investigaci\'{o}n y de Estudios Avanzados (CINVESTAV), Mexico City and M\'{e}rida, Mexico
\item \Idef{org12}Centro Fermi - Museo Storico della Fisica e Centro Studi e Ricerche ``Enrico Fermi'', Rome, Italy
\item \Idef{org13}Chicago State University, Chicago, Illinois, USA
\item \Idef{org14}China Institute of Atomic Energy, Beijing, China
\item \Idef{org15}Commissariat \`{a} l'Energie Atomique, IRFU, Saclay, France
\item \Idef{org16}COMSATS Institute of Information Technology (CIIT), Islamabad, Pakistan
\item \Idef{org17}Departamento de F\'{\i}sica de Part\'{\i}culas and IGFAE, Universidad de Santiago de Compostela, Santiago de Compostela, Spain
\item \Idef{org18}Department of Physics and Technology, University of Bergen, Bergen, Norway
\item \Idef{org19}Department of Physics, Aligarh Muslim University, Aligarh, India
\item \Idef{org20}Department of Physics, Ohio State University, Columbus, Ohio, United States
\item \Idef{org21}Department of Physics, Sejong University, Seoul, South Korea
\item \Idef{org22}Department of Physics, University of Oslo, Oslo, Norway
\item \Idef{org23}Dipartimento di Elettrotecnica ed Elettronica del Politecnico, Bari, Italy
\item \Idef{org24}Dipartimento di Fisica dell'Universit\`{a} 'La Sapienza' and Sezione INFN Rome, Italy
\item \Idef{org25}Dipartimento di Fisica dell'Universit\`{a} and Sezione INFN, Cagliari, Italy
\item \Idef{org26}Dipartimento di Fisica dell'Universit\`{a} and Sezione INFN, Trieste, Italy
\item \Idef{org27}Dipartimento di Fisica dell'Universit\`{a} and Sezione INFN, Turin, Italy
\item \Idef{org28}Dipartimento di Fisica e Astronomia dell'Universit\`{a} and Sezione INFN, Bologna, Italy
\item \Idef{org29}Dipartimento di Fisica e Astronomia dell'Universit\`{a} and Sezione INFN, Catania, Italy
\item \Idef{org30}Dipartimento di Fisica e Astronomia dell'Universit\`{a} and Sezione INFN, Padova, Italy
\item \Idef{org31}Dipartimento di Fisica `E.R.~Caianiello' dell'Universit\`{a} and Gruppo Collegato INFN, Salerno, Italy
\item \Idef{org32}Dipartimento di Scienze e Innovazione Tecnologica dell'Universit\`{a} del  Piemonte Orientale and Gruppo Collegato INFN, Alessandria, Italy
\item \Idef{org33}Dipartimento Interateneo di Fisica `M.~Merlin' and Sezione INFN, Bari, Italy
\item \Idef{org34}Division of Experimental High Energy Physics, University of Lund, Lund, Sweden
\item \Idef{org35}Eberhard Karls Universit\"{a}t T\"{u}bingen, T\"{u}bingen, Germany
\item \Idef{org36}European Organization for Nuclear Research (CERN), Geneva, Switzerland
\item \Idef{org37}Faculty of Engineering, Bergen University College, Bergen, Norway
\item \Idef{org38}Faculty of Mathematics, Physics and Informatics, Comenius University, Bratislava, Slovakia
\item \Idef{org39}Faculty of Nuclear Sciences and Physical Engineering, Czech Technical University in Prague, Prague, Czech Republic
\item \Idef{org40}Faculty of Science, P.J.~\v{S}af\'{a}rik University, Ko\v{s}ice, Slovakia
\item \Idef{org41}Faculty of Technology, Buskerud and Vestfold University College, Vestfold, Norway
\item \Idef{org42}Frankfurt Institute for Advanced Studies, Johann Wolfgang Goethe-Universit\"{a}t Frankfurt, Frankfurt, Germany
\item \Idef{org43}Gangneung-Wonju National University, Gangneung, South Korea
\item \Idef{org44}Gauhati University, Department of Physics, Guwahati, India
\item \Idef{org45}Helsinki Institute of Physics (HIP), Helsinki, Finland
\item \Idef{org46}Hiroshima University, Hiroshima, Japan
\item \Idef{org47}Indian Institute of Technology Bombay (IIT), Mumbai, India
\item \Idef{org48}Indian Institute of Technology Indore, Indore (IITI), India
\item \Idef{org49}Inha University, Incheon, South Korea
\item \Idef{org50}Institut de Physique Nucl\'eaire d'Orsay (IPNO), Universit\'e Paris-Sud, CNRS-IN2P3, Orsay, France
\item \Idef{org51}Institut f\"{u}r Informatik, Johann Wolfgang Goethe-Universit\"{a}t Frankfurt, Frankfurt, Germany
\item \Idef{org52}Institut f\"{u}r Kernphysik, Johann Wolfgang Goethe-Universit\"{a}t Frankfurt, Frankfurt, Germany
\item \Idef{org53}Institut f\"{u}r Kernphysik, Westf\"{a}lische Wilhelms-Universit\"{a}t M\"{u}nster, M\"{u}nster, Germany
\item \Idef{org54}Institut Pluridisciplinaire Hubert Curien (IPHC), Universit\'{e} de Strasbourg, CNRS-IN2P3, Strasbourg, France
\item \Idef{org55}Institute for Nuclear Research, Academy of Sciences, Moscow, Russia
\item \Idef{org56}Institute for Subatomic Physics of Utrecht University, Utrecht, Netherlands
\item \Idef{org57}Institute for Theoretical and Experimental Physics, Moscow, Russia
\item \Idef{org58}Institute of Experimental Physics, Slovak Academy of Sciences, Ko\v{s}ice, Slovakia
\item \Idef{org59}Institute of Physics, Academy of Sciences of the Czech Republic, Prague, Czech Republic
\item \Idef{org60}Institute of Physics, Bhubaneswar, India
\item \Idef{org61}Institute of Space Science (ISS), Bucharest, Romania
\item \Idef{org62}Instituto de Ciencias Nucleares, Universidad Nacional Aut\'{o}noma de M\'{e}xico, Mexico City, Mexico
\item \Idef{org63}Instituto de F\'{\i}sica, Universidad Nacional Aut\'{o}noma de M\'{e}xico, Mexico City, Mexico
\item \Idef{org64}iThemba LABS, National Research Foundation, Somerset West, South Africa
\item \Idef{org65}Joint Institute for Nuclear Research (JINR), Dubna, Russia
\item \Idef{org66}Konkuk University, Seoul, South Korea
\item \Idef{org67}Korea Institute of Science and Technology Information, Daejeon, South Korea
\item \Idef{org68}KTO Karatay University, Konya, Turkey
\item \Idef{org69}Laboratoire de Physique Corpusculaire (LPC), Clermont Universit\'{e}, Universit\'{e} Blaise Pascal, CNRS--IN2P3, Clermont-Ferrand, France
\item \Idef{org70}Laboratoire de Physique Subatomique et de Cosmologie, Universit\'{e} Grenoble-Alpes, CNRS-IN2P3, Grenoble, France
\item \Idef{org71}Laboratori Nazionali di Frascati, INFN, Frascati, Italy
\item \Idef{org72}Laboratori Nazionali di Legnaro, INFN, Legnaro, Italy
\item \Idef{org73}Lawrence Berkeley National Laboratory, Berkeley, California, United States
\item \Idef{org74}Lawrence Livermore National Laboratory, Livermore, California, United States
\item \Idef{org75}Moscow Engineering Physics Institute, Moscow, Russia
\item \Idef{org76}National Centre for Nuclear Studies, Warsaw, Poland
\item \Idef{org77}National Institute for Physics and Nuclear Engineering, Bucharest, Romania
\item \Idef{org78}National Institute of Science Education and Research, Bhubaneswar, India
\item \Idef{org79}Niels Bohr Institute, University of Copenhagen, Copenhagen, Denmark
\item \Idef{org80}Nikhef, National Institute for Subatomic Physics, Amsterdam, Netherlands
\item \Idef{org81}Nuclear Physics Group, STFC Daresbury Laboratory, Daresbury, United Kingdom
\item \Idef{org82}Nuclear Physics Institute, Academy of Sciences of the Czech Republic, \v{R}e\v{z} u Prahy, Czech Republic
\item \Idef{org83}Oak Ridge National Laboratory, Oak Ridge, Tennessee, United States
\item \Idef{org84}Petersburg Nuclear Physics Institute, Gatchina, Russia
\item \Idef{org85}Physics Department, Creighton University, Omaha, Nebraska, United States
\item \Idef{org86}Physics Department, Panjab University, Chandigarh, India
\item \Idef{org87}Physics Department, University of Athens, Athens, Greece
\item \Idef{org88}Physics Department, University of Cape Town, Cape Town, South Africa
\item \Idef{org89}Physics Department, University of Jammu, Jammu, India
\item \Idef{org90}Physics Department, University of Rajasthan, Jaipur, India
\item \Idef{org91}Physik Department, Technische Universit\"{a}t M\"{u}nchen, Munich, Germany
\item \Idef{org92}Physikalisches Institut, Ruprecht-Karls-Universit\"{a}t Heidelberg, Heidelberg, Germany
\item \Idef{org93}Politecnico di Torino, Turin, Italy
\item \Idef{org94}Purdue University, West Lafayette, Indiana, United States
\item \Idef{org95}Pusan National University, Pusan, South Korea
\item \Idef{org96}Research Division and ExtreMe Matter Institute EMMI, GSI Helmholtzzentrum f\"ur Schwerionenforschung, Darmstadt, Germany
\item \Idef{org97}Rudjer Bo\v{s}kovi\'{c} Institute, Zagreb, Croatia
\item \Idef{org98}Russian Federal Nuclear Center (VNIIEF), Sarov, Russia
\item \Idef{org99}Russian Research Centre Kurchatov Institute, Moscow, Russia
\item \Idef{org100}Saha Institute of Nuclear Physics, Kolkata, India
\item \Idef{org101}School of Physics and Astronomy, University of Birmingham, Birmingham, United Kingdom
\item \Idef{org102}Secci\'{o}n F\'{\i}sica, Departamento de Ciencias, Pontificia Universidad Cat\'{o}lica del Per\'{u}, Lima, Peru
\item \Idef{org103}Sezione INFN, Bari, Italy
\item \Idef{org104}Sezione INFN, Bologna, Italy
\item \Idef{org105}Sezione INFN, Cagliari, Italy
\item \Idef{org106}Sezione INFN, Catania, Italy
\item \Idef{org107}Sezione INFN, Padova, Italy
\item \Idef{org108}Sezione INFN, Rome, Italy
\item \Idef{org109}Sezione INFN, Trieste, Italy
\item \Idef{org110}Sezione INFN, Turin, Italy
\item \Idef{org111}SSC IHEP of NRC Kurchatov institute, Protvino, Russia
\item \Idef{org112}SUBATECH, Ecole des Mines de Nantes, Universit\'{e} de Nantes, CNRS-IN2P3, Nantes, France
\item \Idef{org113}Suranaree University of Technology, Nakhon Ratchasima, Thailand
\item \Idef{org114}Technical University of Split FESB, Split, Croatia
\item \Idef{org115}The Henryk Niewodniczanski Institute of Nuclear Physics, Polish Academy of Sciences, Cracow, Poland
\item \Idef{org116}The University of Texas at Austin, Physics Department, Austin, Texas, USA
\item \Idef{org117}Universidad Aut\'{o}noma de Sinaloa, Culiac\'{a}n, Mexico
\item \Idef{org118}Universidade de S\~{a}o Paulo (USP), S\~{a}o Paulo, Brazil
\item \Idef{org119}Universidade Estadual de Campinas (UNICAMP), Campinas, Brazil
\item \Idef{org120}University of Houston, Houston, Texas, United States
\item \Idef{org121}University of Jyv\"{a}skyl\"{a}, Jyv\"{a}skyl\"{a}, Finland
\item \Idef{org122}University of Liverpool, Liverpool, United Kingdom
\item \Idef{org123}University of Tennessee, Knoxville, Tennessee, United States
\item \Idef{org124}University of the Witwatersrand, Johannesburg, South Africa
\item \Idef{org125}University of Tokyo, Tokyo, Japan
\item \Idef{org126}University of Tsukuba, Tsukuba, Japan
\item \Idef{org127}University of Zagreb, Zagreb, Croatia
\item \Idef{org128}Universit\'{e} de Lyon, Universit\'{e} Lyon 1, CNRS/IN2P3, IPN-Lyon, Villeurbanne, France
\item \Idef{org129}V.~Fock Institute for Physics, St. Petersburg State University, St. Petersburg, Russia
\item \Idef{org130}Variable Energy Cyclotron Centre, Kolkata, India
\item \Idef{org131}Vin\v{c}a Institute of Nuclear Sciences, Belgrade, Serbia
\item \Idef{org132}Warsaw University of Technology, Warsaw, Poland
\item \Idef{org133}Wayne State University, Detroit, Michigan, United States
\item \Idef{org134}Wigner Research Centre for Physics, Hungarian Academy of Sciences, Budapest, Hungary
\item \Idef{org135}Yale University, New Haven, Connecticut, United States
\item \Idef{org136}Yonsei University, Seoul, South Korea
\item \Idef{org137}Zentrum f\"{u}r Technologietransfer und Telekommunikation (ZTT), Fachhochschule Worms, Worms, Germany
\end{Authlist}
\endgroup