### 2016-10-27 acknowledgements_jan2015.tex

The ALICE Collaboration would like to thank all its engineers and technicians for their invaluable contributions to the construction of the experiment and the CERN accelerator teams for the outstanding performance of the LHC complex.
%
The ALICE Collaboration gratefully acknowledges the resources and support provided by all Grid centres and the Worldwide LHC Computing Grid (WLCG) collaboration.
%
The ALICE Collaboration acknowledges the following funding agencies for their support in building and
running the ALICE detector:
%
State Committee of Science,  World Federation of Scientists (WFS)
and Swiss Fonds Kidagan, Armenia,
%
Conselho Nacional de Desenvolvimento Cient\'{\i}fico e Tecnol\'{o}gico (CNPq), Financiadora de Estudos e Projetos (FINEP),
Funda\c{c}\~{a}o de Amparo \`{a} Pesquisa do Estado de S\~{a}o Paulo (FAPESP);
%
National Natural Science Foundation of China (NSFC), the Chinese Ministry of Education (CMOE)
and the Ministry of Science and Technology of China (MSTC);
%
Ministry of Education and Youth of the Czech Republic;
%
Danish Natural Science Research Council, the Carlsberg Foundation and the Danish National Research Foundation;
%
The European Research Council under the European Community's Seventh Framework Programme;
%
Helsinki Institute of Physics and the Academy of Finland;
%
French CNRS-IN2P3, the `Region Pays de Loire', `Region Alsace', `Region Auvergne' and CEA, France;
%
German Bundesministerium fur Bildung, Wissenschaft, Forschung und Technologie (BMBF) and the Helmholtz Association;
%
General Secretariat for Research and Technology, Ministry of
Development, Greece;
%
Hungarian Orszagos Tudomanyos Kutatasi Alappgrammok (OTKA) and National Office for Research and Technology (NKTH);
%
Department of Atomic Energy and Department of Science and Technology of the Government of India;
%
Istituto Nazionale di Fisica Nucleare (INFN) and Centro Fermi -
Museo Storico della Fisica e Centro Studi e Ricerche "Enrico
Fermi", Italy;
%
MEXT Grant-in-Aid for Specially Promoted Research, Ja\-pan;
%
Joint Institute for Nuclear Research, Dubna;
%
National Research Foundation of Korea (NRF);
%
Consejo Nacional de Cienca y Tecnologia (CONACYT), Direccion General de Asuntos del Personal Academico(DGAPA), M\'{e}xico, :Amerique Latine Formation academique – European Commission(ALFA-EC) and the EPLANET Program
(European Particle Physics Latin American Network)
%
Stichting voor Fundamenteel Onderzoek der Materie (FOM) and the Nederlandse Organisatie voor Wetenschappelijk Onderzoek (NWO), Netherlands;
%
Research Council of Norway (NFR);
%
National Science Centre, Poland;
%
Ministry of National Education/Institute for Atomic Physics and Consiliul Naţional al Cercetării Ştiinţifice - Executive Agency for Higher Education Research Development and Innovation Funding (CNCS-UEFISCDI) - Romania;
%
Ministry of Education and Science of Russian Federation, Russian
Academy of Sciences, Russian Federal Agency of Atomic Energy,
Russian Federal Agency for Science and Innovations and The Russian
Foundation for Basic Research;
%
Ministry of Education of Slovakia;
%
Department of Science and Technology, South Africa;
%
Centro de Investigaciones Energeticas, Medioambientales y Tecnologicas (CIEMAT), E-Infrastructure shared between Europe and Latin America (EELA), Ministerio de Econom\'{i}a y Competitividad (MINECO) of Spain, Xunta de Galicia (Conseller\'{\i}a de Educaci\'{o}n),
Centro de Aplicaciones Tecnológicas y Desarrollo Nuclear (CEA\-DEN), Cubaenerg\'{\i}a, Cuba, and IAEA (International Atomic Energy Agency);
%
Swedish Research Council (VR) and Knut $\&$ Alice Wallenberg
Foundation (KAW);
%
Ukraine Ministry of Education and Science;
%
United Kingdom Science and Technology Facilities Council (STFC);
%
The United States Department of Energy, the United States National
Science Foundation, the State of Texas, and the State of Ohio;
%
Ministry of Science, Education and Sports of Croatia and  Unity through Knowledge Fund, Croatia.
%
Council of Scientific and Industrial Research (CSIR), New Delhi, India