### 2016-12-19 abstract.tex
first half llya's comment

The correlations between event-by-event fluctuations of anisotropic flow harmonic amplitudes
have been measured in $\PbPb$ collisions at $\snn=2.76$~TeV with the ALICE detector at the Large Hadron Collider. 
The results were obtained with the multiparticle correlation observables dubbed Symmetric Cumulants.
These observables are robust against biases originating from non-flow effects. 
The centrality dependence of correlations between the higher order harmonics (the quadrangular $v_4$ and pentagonal $v_5$ flow) and the lower order harmonics (the elliptic $v_2$ and triangular $v_3$ flow)
is presented. The transverse momentum dependence of correlations between $v_3$ and $v_2$ and between $v_4$ and $v_2$ is also reported. 
The results are compared to calculations from viscous hydrodynamics and  A Multi-Phase Transport ({AMPT}) model calculations.
The comparisons to viscous hydrodynamic models demonstrate that
the different order harmonic correlations respond differently to the initial conditions and the temperature dependence of the ratio of shear viscosity to entropy density ($\eta/s$). 
A small average value of $\eta/s$ is favored independent of specific choice of initial conditions in the models. The calculations with the AMPT initial conditions yield results closest to the measurements. 
Correlations between the magnitudes of $v_2$, $v_3$ and $v_4$ show moderate $p_{\rm T}$ dependence in mid-central collisions. This might be an indication of possible viscous corrections to the equilibrium distribution at hadronic freeze-out, which might help to understand the possible contribution of bulk viscosity in the hadronic phase of the system.
Together with existing measurements of individual flow harmonics, the presented results provide further constraints 
on initial conditions and the transport properties of the system produced in heavy-ion collisions.