### 2017-08-14 appendix.tex
add chisq test

% !TEX root = paper.tex
\appendix
\section{Model Comparisons of the Individual Flow Harmonics $v_n$}
\label{sec:vn}
As discussed in Sec.~\ref{sec:method}, NSC$(m,n)$ is expected to be insensitive to the magnitudes of $v_{m}$ and $v_{n}$ but SC$(m,n)$ has contributions from both the correlations between the two different flow harmonics and the individual harmonics $v_{n}$. Therefore, it is important to check how well the theoretical models used in Sec.~\ref{sec:theory} describe the measured $v_n$ data shown in Sec.~\ref{sec:results}. The comparisons are made only up to $v_4$ because model calculations are not available for $v_5$ at this moment.

\begin{figure*}[h]
            \begin{center}
                       \resizebox{0.34\textwidth}{!}{\includegraphics{figs/FigA1_v2_comp_EKRT.eps}}\hspace{-0.27cm}
                       \resizebox{0.34\textwidth}{!}{\includegraphics{figs/FigA1_v3_comp_EKRT.eps}}\hspace{-0.27cm}
                       \resizebox{0.34\textwidth}{!}{\includegraphics{figs/FigA1_v4_comp_EKRT.eps}}
        \caption{The individual flow harmonics $v_n$ ($n$ = 2, 3 and 4) in $\PbPb$ collisions at $\snn=2.76$~TeV~\cite{Adam:2016izf} are compared to the event-by-event EKRT+viscous hydrodynamic calculations~\cite{Niemi:2015qia}. The lines are hydrodynamic predictions with two different $\eta/s(T)$ parameterizations, labeled in the same way as in ~\cite{Niemi:2015qia}.}
        \label{fig:Figure_A1}
              \end{center}
\end{figure*}

The measured $v_n$ ($n$ = 2, 3 and 4) in $\PbPb$ collisions at $\snn=2.76$~TeV are compared to the event-by-event EKRT+viscous hydrodynamic calculations~\cite{Niemi:2015qia} in Fig.~\ref{fig:Figure_A1}. In these calculations the initial conditions and $\eta/s$ parameterizations are chosen to reproduce the LHC $v_n$ data.
The calculations capture the centrality dependence of $v_n$ in the central and mid-central collisions within 5\% for $v_2$ and 10\% for $v_3$ and $v_4$.

\begin{figure*}[h]
            \begin{center}
                       \resizebox{0.34\textwidth}{!}{\includegraphics{figs/FigA2_v2_comp_VISH.eps}}\hspace{-0.27cm}
                       \resizebox{0.34\textwidth}{!}{\includegraphics{figs/FigA2_v3_comp_VISH.eps}}\hspace{-0.27cm}
                       \resizebox{0.34\textwidth}{!}{\includegraphics{figs/FigA2_v4_comp_VISH.eps}}
        \caption{The individual flow harmonics $v_n$ ($n$ = 2, 3 and 4) in $\PbPb$ collisions at $\snn=2.76$~TeV~\cite{Adam:2016izf} are compared to various VISH2+1 calculations~\cite{Zhu:2016puf}. Three initial conditions from AMPT, MC-KLN and MC-Glauber are drawn as different colors. The $\eta/s$ parameters are shown as different line styles, the small shear viscosity ($\eta/s = 0.08$) are shown as solid lines, and large shear viscosities ($\eta/s = 0.2$ for MC-KLN and MC-Glauber, 0.16 for AMPT) are drawn as dashed lines.}
        \label{fig:Figure_A2}
              \end{center}
\end{figure*}

The VISH2+1 calculations with various initial conditions and $\eta/s$ parameters are compared to the $v_n$ data in Fig.~\ref{fig:Figure_A2}.
Neither MC-Glauber nor MC-KLN initial conditions can simultaneously describe $v_2$, $v_3$ and $v_4$. In particular, for MC-Glauber initial conditions, VISH2+1 with $\eta/s$ = 0.08 can describe well $v_2$ from central to mid-central collisions, but overestimates $v_3$ and $v_4$ for the same centrality ranges. For MC-KLN initial conditions, VISH2+1 with $\eta/s$ = 0.20 reproduces $v_2$ but underestimates $v_3$ and $v_4$ for the presented centrality regions. The calculations with AMPT initial conditions improves the simultaneous descriptions of $v_n$ ($n$ = 2, 3 and 4). The overall difference to the data is quite large if all the model settings are considered, about 30\% for $v_n$ ($n$ = 2 and 3) and 50\% for $v_4$. The calculations with AMPT initial conditions reproduce the observed centrality dependence with an accuracy of 10--20\%.

\begin{figure*}[h]
            \begin{center}
                       \resizebox{0.34\textwidth}{!}{\includegraphics{figs/FigA3_v2_comp_AMPT.eps}}\hspace{-0.27cm}
                       \resizebox{0.34\textwidth}{!}{\includegraphics{figs/FigA3_v3_comp_AMPT.eps}}\hspace{-0.27cm}
                       \resizebox{0.34\textwidth}{!}{\includegraphics{figs/FigA3_v4_comp_AMPT.eps}}
        \caption{The individual flow harmonics $v_n$ ($n$ = 2, 3 and 4) in $\PbPb$ collisions at $\snn=2.76$~TeV~\cite{Adam:2016izf} are compared to various AMPT models.}
        \label{fig:Figure_A3}
              \end{center}
\end{figure*}

The AMPT calculations with various configurations are compared to the $v_n$ data in Fig.~\ref{fig:Figure_A3}.
The string melting version of AMPT~\cite{Lin:2001zk,Lin:2004en} reasonably reproduces $v_n$ as shown in Fig.~\ref{fig:Figure_A3} within 20\% for $v_2$ and 10\% for $v_3$ and $v_4$. The version based on the string melting configuration without the hadronic rescattering phase underestimates the data compared to the calculations with the string melting version of AMPT, which demonstrates that a large fraction of the flow is developed during the late hadronic rescattering stage in the string melting version of AMPT.
The default version of AMPT underestimates $v_n$ ($n$ = 2, 3 and 4) by $\approx$ 20\%. It should be noted that the default AMPT model can describe the normalized symmetric cumulants (NSC($m$,$n$)) quantitively for most centralities while the string melting AMPT model fails to describe them.

\begin{figure*}[h]
            \begin{center}
                       \resizebox{0.34\textwidth}{!}{\includegraphics{figs/FigA4_v2_modelcomparisonBest.eps}}\hspace{-0.27cm}
                       \resizebox{0.34\textwidth}{!}{\includegraphics{figs/FigA4_v3_modelcomparisonBest.eps}}\hspace{-0.27cm}
                       \resizebox{0.34\textwidth}{!}{\includegraphics{figs/FigA4_v4_modelcomparisonBest.eps}}
        \caption{The individual flow harmonics $v_n$ ($n$ = 2, 3 and 4) in $\PbPb$ collisions at $\snn=2.76$~TeV~\cite{Adam:2016izf} are compared to few selected model calculations from three theoretical models which describe the $v_n$ data best.}
        \label{fig:Figure_A4}
              \end{center}
\end{figure*}

Finally, few selected calculations from three theoretical models which describe the $v_n$ data best are shown in Fig.~\ref{fig:Figure_A4}.
The calculations from event-by-event EKRT+viscous hydrodynamics, VISH2+1 with AMPT initial conditions ($\eta/s$ = 0.08) and the string melting version of AMPT give the best description of the individual flow harmonics $v_n$ ($n$ = 2, 3 and 4) with an accuracy of 5--20\%. The centrality dependence differs in the three models as well as in the different order flow harmonics.
Together with SC$(m,n)$ and NSC$(m,n)$, the simultaneous description of individual flow harmonics $v_n$ at all orders is necessary to further optimize model parameters and put better constraints on the initial conditions and the transport properties of nuclear matter in ultra-relativistic heavy-ion collisions.

\newpage
\section{Chi-Square test for SC observables}
\label{sec:chisq}
We perform a least-square test to the models for each SC observable via a log-likelihood or least squares-${\chi}^2$ method~\cite{PDBook} as following,

\begin{equation}
\chi^{2} = \sum_{i=1}^{N} \frac{{(y_i-f_i)}^{2}}{{\sigma_i}^{2}},
\label{Eq:chisq}
\end{equation}

where $y_i$ ($f_i$) is a measurement (model) in a centrality bin, and the systematic and statistical errors  are combined in quadrature $\sigma_i = \sqrt{\sigma_{i, stat}^2 + \sigma_{i, syst}^2}$.
The centrality range is restricted in 10--50\% bins, corresponding the total number of data sets in each observable (N) is 4.
The resulting ${\chi}^2$ for few selected calculations from three theoretical models which describe the SC observables best is shown in Fig.~\ref{fig:Figure_A5}.
\begin{figure*}[h]
            \begin{center}
                       \resizebox{0.8\textwidth}{!}{\includegraphics{figs/chisq_bestfits.eps}}
        \caption{The ${\chi}^2$ values from the comparison of few selected calculations from three theoretical models are shown for all SC observables. Closed (open) markers are for SC($m$,$n$) (NSC($m$,$n$)).}
        \label{fig:Figure_A5}
              \end{center}
\end{figure*}

We found that in case of VISH2+1 calculations, the ${\chi}^2$ values of SC($m$,$n$) is larger than those of NSC($m$,$n$). This can be explained by the fact that the individual flow harmonics $v_n$ is not well described as shown in Fig.~\ref{fig:Figure_A4} compared to other models.
Even though the calculations from event-by-event EKRT+viscous hydrodynamics give best ${\chi}^2$ values both for SC($m$,$n$) and NSC($m$,$n$), the ${\chi}^2$
values are bit large especially for the observables related with $v_3$. 
Even with the best model calculations from three theoretical models, the ${\chi}^2$ value varies a lot depending both on the model parameters and/or different order SC observables, which implies that the different order harmonic correlations have different sensitivities to the initial conditions and the system properties.