### 2017-08-29 fa_2017-08-29.tex
final version submitted, via constantin

% Version: 2017-08-29

The ALICE Collaboration would like to thank all its engineers and technicians for their invaluable contributions to the construction of the experiment and the CERN accelerator teams for the outstanding performance of the LHC complex.
%
The ALICE Collaboration gratefully acknowledges the resources and support provided by all Grid centres and the Worldwide LHC Computing Grid (WLCG) collaboration.
%
The ALICE Collaboration acknowledges the following funding agencies for their support in building and running the ALICE detector:
%
A. I. Alikhanyan National Science Laboratory (Yerevan Physics Institute) Foundation (ANSL), State Committee of Science and World Federation of Scientists (WFS), Armenia;
%
Austrian Academy of Sciences and Nationalstiftung f\"{u}r Forschung, Technologie und Entwicklung, Austria;
%
Ministry of Communications and High Technologies, National Nuclear Research Center, Azerbaijan;
%
Conselho Nacional de Desenvolvimento Cient\'{\i}fico e Tecnol\'{o}gico (CNPq), Universidade Federal do Rio Grande do Sul (UFRGS), Financiadora de Estudos e Projetos (Finep) and Funda\c{c}\~{a}o de Amparo \`{a} Pesquisa do Estado de S\~{a}o Paulo (FAPESP), Brazil;
%
Ministry of Science \& Technology of China (MSTC), National Natural Science Foundation of China (NSFC) and Ministry of Education of China (MOEC) , China;
%
Ministry of Science, Education and Sport and Croatian Science Foundation, Croatia;
%
Ministry of Education, Youth and Sports of the Czech Republic, Czech Republic;
%
The Danish Council for Independent Research | Natural Sciences, the Carlsberg Foundation and Danish National Research Foundation (DNRF), Denmark;
%
Helsinki Institute of Physics (HIP), Finland;
%
Commissariat \`{a} l'Energie Atomique (CEA) and Institut National de Physique Nucl\'{e}aire et de Physique des Particules (IN2P3) and Centre National de la Recherche Scientifique (CNRS), France;
%
Bundesministerium f\"{u}r Bildung, Wissenschaft, Forschung und Technologie (BMBF) and GSI Helmholtzzentrum f\"{u}r Schwerionenforschung GmbH, Germany;
%
General Secretariat for Research and Technology, Ministry of Education, Research and Religions, Greece;
%
National Research, Development and Innovation Office, Hungary;
%
Department of Atomic Energy Government of India (DAE) and Council of Scientific and Industrial Research (CSIR), New Delhi, India;
%
Indonesian Institute of Science, Indonesia;
%
Centro Fermi - Museo Storico della Fisica e Centro Studi e Ricerche Enrico Fermi and Istituto Nazionale di Fisica Nucleare (INFN), Italy;
%
Institute for Innovative Science and Technology , Nagasaki Institute of Applied Science (IIST), Japan Society for the Promotion of Science (JSPS) KAKENHI and Japanese Ministry of Education, Culture, Sports, Science and Technology (MEXT), Japan;
%
Consejo Nacional de Ciencia (CONACYT) y Tecnolog\'{i}a, through Fondo de Cooperaci\'{o}n Internacional en Ciencia y Tecnolog\'{i}a (FONCICYT) and Direcci\'{o}n General de Asuntos del Personal Academico (DGAPA), Mexico;
%
Nederlandse Organisatie voor Wetenschappelijk Onderzoek (NWO), Netherlands;
%
The Research Council of Norway, Norway;
%
Commission on Science and Technology for Sustainable Development in the South (COMSATS), Pakistan;
%
Pontificia Universidad Cat\'{o}lica del Per\'{u}, Peru;
%
Ministry of Science and Higher Education and National Science Centre, Poland;
%
Korea Institute of Science and Technology Information and National Research Foundation of Korea (NRF), Republic of Korea;
%
Ministry of Education and Scientific Research, Institute of Atomic Physics and Romanian National Agency for Science, Technology and Innovation, Romania;
%
Joint Institute for Nuclear Research (JINR), Ministry of Education and Science of the Russian Federation and National Research Centre Kurchatov Institute, Russia;
%
Ministry of Education, Science, Research and Sport of the Slovak Republic, Slovakia;
%
National Research Foundation of South Africa, South Africa;
%
Centro de Aplicaciones Tecnol\'{o}gicas y Desarrollo Nuclear (CEADEN), Cubaenerg\'{\i}a, Cuba, Ministerio de Ciencia e Innovacion and Centro de Investigaciones Energ\'{e}ticas, Medioambientales y Tecnol\'{o}gicas (CIEMAT), Spain;
%
Swedish Research Council (VR) and Knut \& Alice Wallenberg Foundation (KAW), Sweden;
%
European Organization for Nuclear Research, Switzerland;
%
National Science and Technology Development Agency (NSDTA), Suranaree University of Technology (SUT) and Office of the Higher Education Commission under NRU project of Thailand, Thailand;
%
Turkish Atomic Energy Agency (TAEK), Turkey;
%
National Academy of  Sciences of Ukraine, Ukraine;
%
Science and Technology Facilities Council (STFC), United Kingdom;
%
National Science Foundation of the United States of America (NSF) and United States Department of Energy, Office of Nuclear Physics (DOE NP), United States of America.